### 2018-05-21 4_impl.tex
Add some results to eval.

The previous version of \porthos{} did not distinguish the event-based program model from the high-level AST, they both were encoded into a single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos[1]).
Moreover, the syntax tree was implemented as a mutable data structure, which is being modified at all stages of the program (for instance, see the methods `\texttt{dartagnan.program.Program.compile(...)}' of \porthos{} that recursively compute some properties of the AST and change its state).
We are inclined to consider this architecture as one that is fast to develop, but hard to maintain (since it is difficult to guarantee the correctness of the program) and extend (since adding the support for a new high-level instruction requires changing multiple components of the program, from parser to encoder).
%Moreover, we encounter serious problems when we need to add support for control-flow jumps (as \texttt{continue}, \texttt{break}, \texttt{goto} in C).

@@

\vspace{0.5em}
\begin{enumerate}[nolistsep]
  \item \textit{Robustness:}
    \begin{enumerate}[label*=\arabic*.]
      \item preservation the completeness of analysis,
      \item modular architecture: each module can be tested independently,
      \item usage of software design patterns where necessary, and
      \item usage of immutable data structures for all DTOs.
    \end{enumerate}
  \item \textit{Transparency:}
    \begin{enumerate}[label*=\arabic*.]
      \item following the principles of simplicity and readability, and
      \item clear and informative program output.
    \end{enumerate}
  \item \textit{Efficiency:}
    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1.5em]
      \item keeping the trade-off between execution time and memory usage.
    \end{enumerate}
  \item \textit{Extensibility:}
    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1em]
      \item clear modular architecture.
    \end{enumerate}
\end{enumerate}

Robustness of the analysis is the main criterion of \porthos[2] as a verification tool.
Although it makes the analysis more sensitive to the combinatorial explosion, it preserves the completeness of analysis, which is necessary for a model-checking tool.
As a robust and transparent tool, \porthos[2] must adhere to the strategy of aborting its work on any unexpected outcome (for instance, if a parser failed to parse the string and the recovery algorithm is not described).
%The modular architecture of \porthos[2] allow each its module to be tested independently.
%Usage of immutable data structures for all DTOs helps 
%immut: preservation of invariants
%The transparency, efficiency and extensibility are the necessary criteria of any good tool.

>>>>

The previous version of \porthos{} did not distinguish the event-based program model from the high-level AST, they both were encoded into a single SMT-formula (see classes of package `\texttt{dartagnan.program}' of \porthos[1]).
%For 
Moreover, the syntax tree was implemented as a mutable data structure, which is being modified at all stages of the program (for instance, see the methods `\texttt{dartagnan.program.Program.compile(...)}' of \porthos{} that recursively compute some properties of the AST and change its state).
We are inclined to consider this architecture as one that is fast to develop, but hard to maintain (since it is difficult to guarantee the correctness of the program) and extend (since adding the support for a new high-level instruction requires changing multiple components of the program, from parser to encoder).
%Moreover, we encounter serious problems when we need to add support for control-flow jumps (as \texttt{continue}, \texttt{break}, \texttt{goto} in C).

@@

\vspace{0.5em}
\begin{enumerate}[nolistsep]
  \item \textit{Robustness:}
    \begin{enumerate}[label*=\arabic*.]
      \item preservation the completeness of analysis,
      \item modular architecture: each module can be tested independently,
      \item usage of software design patterns where necessary, and
      \item usage of immutable data structures for all DTOs.
    \end{enumerate}
  \item \textit{Transparency:}
    \begin{enumerate}[label*=\arabic*.]
      \item following the principles of simplicity and readability,
      \item clear and informative program output, and
      \item following the clear code style.
    \end{enumerate}
  \item \textit{Efficiency:}
    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1.5em]
      \item keeping the trade-off between execution time and memory usage.
    \end{enumerate}
  \item \textit{Extensibility:}
    \begin{enumerate}[label*=\arabic*.]%[leftmargin=1em]
      \item clear modular architecture.
    \end{enumerate}
\end{enumerate}

Robustness of the analysis is the main criterion of \porthos[2] as a verification tool.
Although it makes the analysis more sensitive to the combinatorial explosion, it preserves the completeness of analysis, which is necessary for a model-checking tool.
As a robust and transparent tool, \porthos[2] must adhere to the strategy of aborting its work on any unexpected outcome (for instance, if a parser failed to parse the string and the recovery algorithm is not described).
One of the transparency principles is following the clear code style.
This mainly means the clear and informative naming of classes, functions and fields.
It also implies the unified ordering of methods in classes, minimised size of methods code, clear tabulation, etc.
%The modular architecture of \porthos[2] allow each its module to be tested independently.
%Usage of immutable data structures for all DTOs helps 
%immut: preservation of invariants
%The transparency, efficiency and extensibility are the necessary criteria of any good tool.

### 2018-05-21 5_eval.tex
Cleanup some old todos.
Add some results to eval.

\begin{figure}[!h]
\includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/test_unrolled.png}
\caption{An example of an unrolled event-flow graph represented in Figure~\ref{ex:compilation}}
\label{ex:unrolling}
\end{figure}

\subsection{Reachability analysis}

Measure time by \porthos[1], \porthos[2]:

- enc-time absolute values

- rate enc-time/solve-time

>>>>

\begin{figure}[!h]
\includegraphics[width=\textwidth,keepaspectratio]{img/my/graphs/test_unrolled.png}
\caption{Unrolling of the event-flow graph presented in Figure~\ref{ex:compilation}}
\label{ex:unrolling}
\end{figure}

\section{Performance}

\subsection{State reachability analysis}

As an example of the C program liable for the reachability and portability analysis, consider the Dekker's algorithm for mutual exclusion of two processes, originally described by Dijkstra~\cite{dijkstra1962over}; the program is presented in Appendix~\ref{apx:dekker}.

For checking the reachability of the final state of the last states (sink events) of unrolled graphs compiled for a certain hardware architecture, \porthos[2] encodes both program and hardware memory model constraints as it was discussed in Section~\ref{ch:enc:bmc}.
%The result formula is then solved by the Z3

%TO-DO: fix interface
%To run \porthos[2] in reachability analysis mode, we use the command `\lstinline{mousquetaires.DartagnanApp -i benchmarks/litmus_C11/Dekker.c -s tso}'

The standard output of the tool looks as following:
\begin{lstlisting}
$ java Dartagnan -i benchmarks/C11/Dekker.c -s tso
Interpreting...
Unrolling...
Program encoding...
Memory model encoding...
#=59
Solving...
{
  "result": "NonReachable",
  "interpretationTimer": {
    "elapsedTimeSec": 0.134
  },
  "unrollingTimer": {
    "elapsedTimeSec": 0.031
  },
  "programEncodingTimer": {
    "elapsedTimeSec": 0.14
  },
  "memoryModelEncodingTimer": {
    "elapsedTimeSec": 2.984
  },
  "solvingTimer": {
    "elapsedTimeSec": 0.037
  },
  "errors": []
}
\end{lstlisting}

For time benchmarking we ran the tool 5 times and computed the median of the encoding time.
Benchmarking was performed on the Linux machine 8-core Intel(R) Core(TM) i7-3632QM CPU @ 2.20GHz, Java(TM) SE Runtime Environment (build 1.8.0\_161-b12) (Java virtual machine was configured by default parameters).
The time was measured by the tool itself via the Java method \texttt{System.currentTimeMillis}.
For \porthos[1] we specify the unrolling bound $k_1 = 2$ (which means that all loops are executed twice), and for \porthos[2] we specify the bound $k_2 = 32$ (which means approximately the same, because

\porthos[2] shows the full encoding time for the Dekker's algorithm 2.699 sec (from which 0.152 sec spent for the program encoding, and other part spent for the memory model encoding).
In contrast, \porthos[1] shows the encoding time 6.152 sec (from which 0.223 sec spent for the program encoding). %, which is considerably better result.
Thus, the performance of the encoding stage has been improved \textbf{in 2.7 times}.

As the time spent for the interpretation and unrolling stages is negligible comparing to the encoding time, we conclude that the new architecture implies minor performance overhead comparing to the previous version of the tool.
The time spent by the SMT solver while considering the SMT formula encoded by \porthos[2] was 0.039 sec, and in case of \porthos[1] the time was 1.291 sec.
The number of events in the event-flow graph of \porthos[2] is 82 events (among them 59 memory events), and the number of events encoded by \porthos[1] is 95 (among them 51 memory events).
Even though the number of memory events processed by \porthos[1] was less, the time of solving the result SMT-formula was significantly more.
We consider this improvement to be the result of technical optimisations applied to \porthos[2].
The optimisations include:
\begin{itemize}
\item \textit{Memoisation} of the frequently requested calls. \\
    For example, the code of \porthos[1] contains 39 calls that take the subset of a set of events. The pattern is the following: `\lstinline{program.getEvents().stream().filter(e -> e instanceof MemEvent || e instanceof Local).collect(Collectors.toSet())}', \\
    10 of them are performed in a loop over all events (in \texttt{Domain.encode} method of \porthos[2]).
    These calls were replaced by the lazily initialisable memoising methods of \texttt{XProgram}: \\
\begin{lstlisting}[language=Java]
public ImmutableSet<XMemoryEvent> getMemoryEvents() {
    return memoryEvents != null
            ? memoryEvents
            : (memoryEvents = getAllNodesExceptSource(XMemoryEvent.class));
}
\end{lstlisting}
\item Usage of \textit{immutable data structures} (as it was discussed in Chapter~\ref{ch:impl}).
\end{itemize}

<TODO: compare numbers of clauses of SMT formulas> !!!

<TODO: profile both programs! And say sth about memory usage!>

% state reachability
